---
title: "First tests of new MAGE model in BaM"
author: "Felipe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
```{r load-packages, message = FALSE, include= FALSE}
library(RBaM)
library(gridExtra)
library(ggplot2)
library(dplyr)
library(tidyr)
```

# New features in RBaM for MAGE model calibration:

The new features in RBaM for the MAGE model calibration include:

- Model with 4 input variables:
  - **X1**: event number starting from 1.
  - **X2**: reach number starting from 1.
  - **X3**: position coordinate in the reach. Units depends on the MAGE projet.
  - **X4**: time coordinate in **seconds**.
- Model with 3 output variables. Units depends on the MAGE projet.:
  - **Y1**: water surface elevation (WSE).
  - **Y2**: discharge.
  - **Y3**: velocity. 
- Calibration process multiple events **(already tested)** and multiple reaches **(not supported yet)**.
- Need to specify several folders for running several events at the same time during the calibration process.
- Spatial-temporal error model specified to each output variable: WSE, discharge and velocity error model.
- Legendre polynomial is already tested for the friction spatially distributed.
- Get simulation data using the `.BIN` file, instead of `mage_extraire`.
- **Limitations:** 
  - The model can receive velocity measurements during calibration, but it is necessary to implement the velocity simulaiton in mage based on wetted area and discharge to get the simulation data.

# Setup for the MAGE model

The simulation is performed with 1200 simulations (30 NAdapt x 40 Ncycles) and 4 events. The data is stored in a text file named `CalibrationData.txt`. The data includes the following columns with -9999 as missing values:

- **Event**: event number.
- **Reach**: reach number.
- **x**: position coordinate in the reach.
- **t**: time coordinate in the reach.
- **Y_Z**: observed water surface elevation (WSE).
- **Y_Q**: observed discharge.
- **Y_V**: observed velocity.
- **Yu_Z**: uncertainty of the observed water surface elevation (WSE).
- **Yu_Q**: uncertainty of the observed discharge.
- **Yu_V**: uncertainty of the observed velocity.


```{r}
path <- "/home/famendezrios/Documents/These/VSCODE-R/HydroBayes/HydroBayes_git/Development/New_features_Ben/"
path_ndegree_1 <- "/home/famendezrios/Documents/These/VSCODE-R/HydroBayes/HydroBayes_git/Development/New_features_Ben_spatial"
# Read calibration data and summary
getCalData <- read.table(
  file = file.path(path, "CalibrationData.txt"),
  header = TRUE,
  stringsAsFactors = FALSE
)
knitr::kable(getCalData,
  align = "c"
)
```

# Case constant friction (degree=0):

## Check of MCMC results 
```{r}
mcmc <- read.table(
  file = file.path(path, "Results_MCMC.txt"),
  header = TRUE,
  stringsAsFactors = FALSE
)

# Trace plot for each parameter
plots <- tracePlot(mcmc)
gridExtra::grid.arrange(grobs = plots, ncol = 3)

# Density plot for each parameter
plots <- densityPlot(mcmc)
gridExtra::grid.arrange(grobs = plots, ncol = 3)
```


## Check summary

```{r}
getSummary <- read.table(
  file = file.path(path, "Results_Summary.txt"),
  header = TRUE,
  stringsAsFactors = FALSE
)

# Values of error model in meter for WSE, in m3/s for discharge and m/s for velocity.
knitr::kable(getSummary,
  align = "c"
)

# Zoom into the MAP and standard deviation of the error model
getSummary_zoom <- getSummary[c(11, 16), ]
getSummary_zoom[, -1] <- getSummary_zoom[, -1] * 1000 # Convert to mm for WSE and LPS for discharge, and mm/s for velocity
# Be careful with discharge and velocity, they are not observation, but only simulated values.
knitr::kable(getSummary_zoom,
  align = "c",
  caption = "Zoom into the MAP and standard deviation of the error model: WSE in mm, discharge in L/s and velocity in mm/s."
)
```

## Check spatial friction distribution

```{r}
n_degree <- 0

ks_literature <- data.frame(min = 1 / 0.013, max = 1 / 0.009, mean = 1 / 0.010)


matrix_zFileKmin <- read.table(file.path(path, "Zfile_Kmin.txt"), header = TRUE)

position <- read.table(file.path(path, "vector_legendre.txt"), header = TRUE)

# Get MAP simulation
MAP_param_matrix <- as.numeric(getSummary_zoom[2, c(1:(n_degree + 1))])

k_estimated_all <- as.data.frame(as.matrix(matrix_zFileKmin) %*% as.matrix(t(mcmc[, 1:(n_degree + 1)])))

k_estimated_all$KP <- position[, 1]

# Convert to long format
df_MCMC_sampling <- pivot_longer(
  k_estimated_all,
  cols = -KP,
  values_to = "Value"
) %>%
  select(KP, Value) %>%
  mutate(ID = "MCMC Sampling")

k_estimated_MAP <- as.matrix(matrix_zFileKmin) %*% MAP_param_matrix

df_MAP <- data.frame(
  KP = position[, 1],
  Value = k_estimated_MAP,
  ID = "MAP"
)

# Get 95% uncertainty for envelope curve : create ribbon data from MCMC
df_envelope <- df_MCMC_sampling %>%
  filter(ID == "MCMC Sampling") %>%
  group_by(KP) %>%
  summarise(
    ymin = quantile(Value, probs = 0.025, na.rm = TRUE),
    ymax = quantile(Value, probs = 0.975, na.rm = TRUE),
    ID = "Parametric\nuncertainty", # so we can map to fill
    .groups = "drop"
  )
ggplot() +
  geom_ribbon(
    data = df_envelope,
    aes(x = KP, ymin = ymin, ymax = ymax, fill = ID)
  ) +
  geom_line(
    data = df_MAP,
    aes(x = KP, y = Value, color = ID)
  ) +
  geom_hline(aes(yintercept = ks_literature$min, linetype = "ASCI (1980)"), color = "gray") +
  geom_hline(aes(yintercept = ks_literature$max, linetype = "ASCI (1980)"), color = "gray") +
  geom_hline(aes(yintercept = ks_literature$mean, linetype = "ASCI (1980)"), color = "gray") +
  scale_fill_manual(values = c("Parametric\nuncertainty" = "pink")) +
  scale_color_manual(values = c("MAP" = "black")) +
  # Échelle de type de ligne
  scale_linetype_manual(
    name = "Reference\nvalues",
    values = c("ASCI (1980)" = "dashed")
  ) +
  labs(
    title = "Friction coefficient estimation \nwith parametric uncertainty",
    x = "Lengthwise position (meters)",
    y = expression("Friction coefficient (m"^
      {
        1 / 3
      } * "/s)"),
    fill = "95% credibility\ninterval",
    color = NULL
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.title = element_text(hjust = 0.5)
  )
```

## Check of residuals

```{r}
# Read residuals
residuals <- read.table(
  file = file.path(path, "Results_Residuals.txt"),
  header = TRUE,
  stringsAsFactors = FALSE
)

# Replace -9999 and -1e9 with NA
residuals[residuals == -9999] <- NA
residuals[residuals == -1e9] <- NA

# Convert residuals to mm and m3/s
residuals_mm_m3_s <- data.frame(residuals[, 1:8], residuals[, -c(1:8)] * 1000)

residuals_event_mm_m3_s <- data.frame(getCalData[, 1:4], residuals_mm_m3_s)

# Plot WSE residuals
ggplot(residuals_event_mm_m3_s, aes(x = X3_obs, shape = factor(Event))) +
  geom_point(aes(y = Y1_sim, col = "sim")) +
  labs(title = "MAP simulations vs Observations (WSE)", x = "Position (m)", y = "Water surface elevation (mm)") +
  geom_point(aes(y = Y1_obs, col = "obs")) +
  scale_color_manual(values = c("sim" = "blue", "obs" = "red")) +
  labs(colour = NULL, shape = "Events") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5), )

# Plot Q residuals
ggplot(residuals_event_mm_m3_s, aes(x = X3_obs, shape = factor(Event))) +
  geom_point(aes(y = Y2_sim, col = "sim")) +
  labs(title = "MAP simulations vs boundary condition in Q", x = "Position (m)", y = "Q (L/s)") +
  geom_hline(aes(yintercept = 120, col = "boundary\nconditions", linetype = "1")) +
  geom_hline(aes(yintercept = 60, col = "boundary\nconditions", linetype = "2")) +
  geom_hline(aes(yintercept = 30, col = "boundary\nconditions", linetype = "3")) +
  geom_hline(aes(yintercept = 14, col = "boundary\nconditions", linetype = "4")) +
  scale_color_manual(values = c("sim" = "blue", "boundary\nconditions" = "red")) +
  scale_linetype_manual(values = c("1" = "dashed", "2" = "solid", "3" = "dotted", "4" = "twodash")) +
  labs(colour = NULL, shape = "Simulated\nevents", linetype = "boundary\ncondition\nevents") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

# Case linear spatial distribution of friction (degree 1):

The calibration data is the same because the only difference is the friction distribution, which is now spatially distributed with a Legendre polynomial of degree 1.

Here the files which changed:

- Config_Model: model has now a new parameter a1_min to be estimated.
- Zfile_Kmin.txt: add a new column which represents the Legendre polynomial of degree 1.
- Config_setup.txt: modify the number of columns in the Zfile_Kmin.txt file to 2.


## Check of MCMC results 
```{r}
mcmc <- read.table(
  file = file.path(path_ndegree_1, "Results_MCMC.txt"),
  header = TRUE,
  stringsAsFactors = FALSE
)

# Trace plot for each parameter
plots <- tracePlot(mcmc)
gridExtra::grid.arrange(grobs = plots, ncol = 3)

# Density plot for each parameter
plots <- densityPlot(mcmc)
gridExtra::grid.arrange(grobs = plots, ncol = 3)
```


## Check summary

```{r}
getSummary <- read.table(
  file = file.path(path_ndegree_1, "Results_Summary.txt"),
  header = TRUE,
  stringsAsFactors = FALSE
)

# Values of error model in meter for WSE, in m3/s for discharge and m/s for velocity.
knitr::kable(getSummary,
  align = "c"
)

# Zoom into the MAP and standard deviation of the error model
getSummary_zoom <- getSummary[c(11, 16), ]
getSummary_zoom[, -c(1, 2)] <- getSummary_zoom[, -c(1, 2)] * 1000 # Convert to mm for WSE and LPS for discharge, and mm/s for velocity
# Be careful with discharge and velocity, they are not observation, but only simulated values.
knitr::kable(getSummary_zoom,
  align = "c",
  caption = "Zoom into the MAP and standard deviation of the error model: WSE in mm, discharge in L/s and velocity in mm/s."
)
```

## Check spatial friction distribution propaging parametric uncertainty

```{r}
n_degree <- 1

matrix_zFileKmin <- read.table(file.path(path_ndegree_1, "Zfile_Kmin.txt"), header = TRUE)

position <- read.table(file.path(path_ndegree_1, "vector_legendre.txt"), header = TRUE)

# Get MAP simulation
MAP_param_matrix <- as.numeric(getSummary_zoom[2, c(1:(n_degree + 1))])

k_estimated_all <- as.data.frame(as.matrix(matrix_zFileKmin) %*% as.matrix(t(mcmc[, 1:(n_degree + 1)])))

k_estimated_all$KP <- position[, 1]

# Convert to long format
df_MCMC_sampling <- pivot_longer(
  k_estimated_all,
  cols = -KP,
  values_to = "Value"
) %>%
  select(KP, Value) %>%
  mutate(ID = "MCMC Sampling")

k_estimated_MAP <- as.matrix(matrix_zFileKmin) %*% MAP_param_matrix

df_MAP <- data.frame(
  KP = position[, 1],
  Value = k_estimated_MAP,
  ID = "MAP"
)

# Get 95% uncertainty for envelope curve : create ribbon data from MCMC
df_envelope <- df_MCMC_sampling %>%
  filter(ID == "MCMC Sampling") %>%
  group_by(KP) %>%
  summarise(
    ymin = quantile(Value, probs = 0.025, na.rm = TRUE),
    ymax = quantile(Value, probs = 0.975, na.rm = TRUE),
    ID = "Parametric\nuncertainty", # so we can map to fill
    .groups = "drop"
  )
ggplot() +
  geom_ribbon(
    data = df_envelope,
    aes(x = KP, ymin = ymin, ymax = ymax, fill = ID)
  ) +
  geom_line(
    data = df_MAP,
    aes(x = KP, y = Value, color = ID)
  ) +
  geom_hline(aes(yintercept = ks_literature$min, linetype = "ASCI (1980)"), color = "gray") +
  geom_hline(aes(yintercept = ks_literature$max, linetype = "ASCI (1980)"), color = "gray") +
  geom_hline(aes(yintercept = ks_literature$mean, linetype = "ASCI (1980)"), color = "gray") +
  scale_fill_manual(values = c("Parametric\nuncertainty" = "pink")) +
  scale_color_manual(values = c("MAP" = "black")) +
  # Échelle de type de ligne
  scale_linetype_manual(
    name = "Reference\nvalues",
    values = c("ASCI (1980)" = "dashed")
  ) +
  labs(
    title = "Friction coefficient estimation \nwith parametric uncertainty",
    x = "Lengthwise position (meters)",
    y = expression("Friction coefficient (m"^
      {
        1 / 3
      } * "/s)"),
    fill = "95% credibility\ninterval",
    color = NULL
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.title = element_text(hjust = 0.5)
  )
```

## Check of residuals

```{r}
# Read residuals
residuals <- read.table(
  file = file.path(path_ndegree_1, "Results_Residuals.txt"),
  header = TRUE,
  stringsAsFactors = FALSE
)

# Replace -9999 and -1e9 with NA
residuals[residuals == -9999] <- NA
residuals[residuals == -1e9] <- NA

# Convert residuals to mm and m3/s
residuals_mm_m3_s <- data.frame(residuals[, 1:8], residuals[, -c(1:8)] * 1000)

residuals_event_mm_m3_s <- data.frame(getCalData[, 1:4], residuals_mm_m3_s)

# Plot WSE residuals
ggplot(residuals_event_mm_m3_s, aes(x = X3_obs, shape = factor(Event))) +
  geom_point(aes(y = Y1_sim, col = "sim")) +
  labs(title = "MAP simulations vs Observations (WSE)", x = "Position (m)", y = "Water surface elevation (mm)") +
  geom_point(aes(y = Y1_obs, col = "obs")) +
  scale_color_manual(values = c("sim" = "blue", "obs" = "red")) +
  labs(colour = NULL, shape = "Events") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5), )

# Plot Q residuals
ggplot(residuals_event_mm_m3_s, aes(x = X3_obs, shape = factor(Event))) +
  geom_point(aes(y = Y2_sim, col = "sim")) +
  labs(title = "MAP simulations vs boundary condition in Q", x = "Position (m)", y = "Q (L/s)") +
  geom_hline(aes(yintercept = 120, col = "boundary\nconditions", linetype = "1")) +
  geom_hline(aes(yintercept = 60, col = "boundary\nconditions", linetype = "2")) +
  geom_hline(aes(yintercept = 30, col = "boundary\nconditions", linetype = "3")) +
  geom_hline(aes(yintercept = 14, col = "boundary\nconditions", linetype = "4")) +
  scale_color_manual(values = c("sim" = "blue", "boundary\nconditions" = "red")) +
  scale_linetype_manual(values = c("1" = "dashed", "2" = "solid", "3" = "dotted", "4" = "twodash")) +
  labs(colour = NULL, shape = "Simulated\nevents", linetype = "boundary\ncondition\nevents") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

# Conclusions

  - The new features are already implemented in BaM! and it is working. 
  - New MAGE model is fixed to 4 inputs (event, reach, x coordinate, t coordinate) and 3 outputs (WSE, discharge, velocity)
  - Time coordinates will always be in seconds, because the `.BIN` file required it.
  - Estimation of friction coefficients looks well and coherent with previous results using boundary conditions with all events merged in a only projet.

# To Do

  - Create new structural error models for WSE, discharge and velocity
  - Create a python module to implement a function to get the initial condition for each event 
  - Finish writing the structural error model's equations until likelihood fonction. 
  - Estimate the model using RBaM functions to calculate the loglikelihood outside BaM and its restrictions.  