---
title: "Checking MCMC Results and Residuals"
author: "Felipe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
```{r load-packages, message = FALSE, include= FALSE}
library(RBaM)
library(gridExtra)
library(ggplot2)
```
# First tests of new mage model in BaM

## New features in RBaM for Mage model calibration:

The new features in RBaM for the Mage model calibration include:

- Model with 4 input variables:
  - **X1**: event number starting from 1.
  - **X2**: reach number starting from 1.
  - **X3**: position coordinate in the reach.
  - **X4**: time coordinate in the reach.
- Model with 3 output variables:
  - **Y1**: water surface elevation (WSE).
  - **Y2**: discharge.
  - **Y3**: velocity. 
- Calibration process multiple events **(being tested)** and multiple reaches **(not supported yet)**.
- Need to specify several folders for running several events at the same time during the calibration process.
- Spatial-temporal error model specified to each output variable: WSE, discharge and velocity error model.
- Legendre polynomial are not tested yet but should work.
- **Limitation:** the error model can not be different for each event, but it can be different for each output variable.

## Setup for the Mage model

The simulation is performed with 625 simulations (25 NAdapt x 25 Ncycles) and 4 events. The data is stored in a text file named `CalibrationData.txt`. The data includes the following columns with -9999 as missing values:

- **Event**: event number.
- **Reach**: reach number.
- **x**: position coordinate in the reach.
- **t**: time coordinate in the reach.
- **Y_Z**: observed water surface elevation (WSE).
- **Y_Q**: observed discharge.
- **Y_V**: observed velocity.
- **Yu_Z**: uncertainty of the observed water surface elevation (WSE).
- **Yu_Q**: uncertainty of the observed discharge.
- **Yu_V**: uncertainty of the observed velocity.


```{r}
path <- "/home/famendezrios/Documents/These/VSCODE-R/HydroBayes/HydroBayes_git/Development/New_features_Ben/"
# Read calibration data and summary
getCalData <- read.table(
  file = file.path(path, "CalibrationData.txt"),
  header = TRUE,
  stringsAsFactors = FALSE
)
knitr::kable(getCalData,
  align = "c"
)
```

## Check of MCMC results 
```{r}
mcmc <- read.table(
  file = file.path(path, "Results_MCMC.txt"),
  header = TRUE,
  stringsAsFactors = FALSE
)

# Trace plot for each parameter
plots <- tracePlot(mcmc)
gridExtra::grid.arrange(grobs = plots, ncol = 3)

# Density plot for each parameter
plots <- densityPlot(mcmc)
gridExtra::grid.arrange(grobs = plots, ncol = 3)
```

## Check summary

```{r}
getSummary <- read.table(
  file = file.path(path, "Results_Summary.txt"),
  header = TRUE,
  stringsAsFactors = FALSE
)

# Values of error model in meter for WSE, in m3/s for discharge and m/s for velocity.
knitr::kable(getSummary,
  align = "c"
)

# Zoom into the MAP and standard deviation of the error model
getSummary_zoom <- getSummary[c(11, 16), ]
getSummary_zoom[, -1] <- getSummary_zoom[, -1] * 1000 # Convert to mm for WSE and LPS for discharge, and mm/s for velocity
# Be careful with discharge and velocity, they are not observation, but only simulated values.
knitr::kable(getSummary_zoom,
  align = "c",
  caption = "Zoom into the MAP and standard deviation of the error model: WSE in mm, discharge in L/s and velocity in mm/s."
)
```


## Check of residuals

```{r}
# Read residuals
residuals <- read.table(
  file = file.path(path, "Results_Residuals.txt"),
  header = TRUE,
  stringsAsFactors = FALSE
)

# Replace -9999 and -1e9 with NA
residuals[residuals == -9999] <- NA
residuals[residuals == -1e9] <- NA

# Convert residuals to mm and m3/s
residuals_mm_m3_s <- data.frame(residuals[, 1:8], residuals[, -c(1:8)] * 1000)

residuals_event_mm_m3_s <- data.frame(getCalData[, 1:4], residuals_mm_m3_s)

# Plot WSE residuals
ggplot(residuals_event_mm_m3_s, aes(x = X3_obs, shape = factor(Event))) +
  geom_point(aes(y = Y1_sim, col = "sim")) +
  labs(title = "MAP simulations vs Observations (WSE)", x = "Position (m)", y = "Water surface elevation (mm)") +
  geom_point(aes(y = Y1_obs, col = "obs")) +
  scale_color_manual(values = c("sim" = "blue", "obs" = "red")) +
  labs(colour = NULL, shape = "Events") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5), )

# Plot Q residuals
ggplot(residuals_event_mm_m3_s, aes(x = X3_obs, shape = factor(Event))) +
  geom_point(aes(y = Y2_sim, col = "sim")) +
  labs(title = "MAP simulations vs boundary condition in Q", x = "Position (m)", y = "Q (L/s)") +
  geom_hline(aes(yintercept = 120, col = "boundary\nconditions", linetype = "1")) +
  geom_hline(aes(yintercept = 60, col = "boundary\nconditions", linetype = "2")) +
  geom_hline(aes(yintercept = 30, col = "boundary\nconditions", linetype = "3")) +
  geom_hline(aes(yintercept = 14, col = "boundary\nconditions", linetype = "4")) +
  scale_color_manual(values = c("sim" = "blue", "boundary\nconditions" = "red")) +
  scale_linetype_manual(values = c("1" = "dashed", "2" = "solid", "3" = "dotted", "4" = "twodash")) +
  labs(colour = NULL, shape = "Simulated\nevents", linetype = "boundary\ncondition\nevents") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```