---
title: "Comparison of calibration methods and polynomial degree"
author: "Felipe"
date: "2025-06-13"
output: html_document
---

# Case study : open-channel at HHlab

## Description
All the information about the case of study could be consulted here : the paper [Proust et al., (2022)](doi:10.1017/jfm.2022.818) and the [Dataset Proust et al., (2022)](https://doi.org/10.57745/EQURJN):

- Rectangular channel : 18 m long, 2 m wide
- Material : glass
- slope : 1.04 x 10-3
- Discharges : 14, 30, 60, 120 L/s (uniform condition smooth bed) 
- Water depth measurements: 161 measures, x = 0.50 m and 16.50 m

Two methods for calibration were implemented, a time series and a water surface profiles calibration. So far, BaM! can calibrate a model only in a single dimension, in other words, either time series nor water surface profiles, but not simultaneously. That's why the exercise was to implement a script to make both of them for using afterwards. 


```{r setup, include=FALSE}
workspace <- "/home/famendezrios/Documents/These/VSCODE-R/HydroBayes/HydroBayes_git/Case_studies/HHLab/"
backup_path <- "Backup"
simulation <- "10_06_2025_sans_influence_aval"

root_dir <- file.path(workspace, backup_path, simulation)

knitr::opts_knit$set(root.dir = root_dir)

# Set consistent figure output options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 5,
  fig.align = "center",
  out.width = "80%"
)
```

```{r}
path_cal_case <- c("Calibration_time_series", "Calibration_water_surface_profiles")
nY <- c()
for (j in path_cal_case) {
  nY_temp <- ifelse(j == "Calibration_time_series", 8, 4)
  nY <- c(nY, nY_temp)
}

path_results <- file.path(backup_path, simulation, path_cal_case)
# Legendre polynomial
n_legendre_polynomial <- 0

path_calibration <- file.path(path_cal_case, paste0("n_", n_legendre_polynomial))

# Load calibration data:
load(file.path(
  "/home/famendezrios/Documents/These/VSCODE-R/HydroBayes/HydroBayes_git/data/processed_data/HHLab/smooth_bed",
  "data_HHLab_all_cases.RData"
))

knitr::kable(head(WS_profiles$case_60_60),
  align = "c", row.names = F
)
```

```{r}
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidyr)

i <- which(path_cal_case == "Calibration_water_surface_profiles")

# Check boundary conditions of 1D model
Upstream_BC <- read.table(file.path(path_calibration[i], "model_mage", "smooth_bed.HYD"),
  header = FALSE,
  skip = 4,
  col.names = c("time", "discharge")
)
Upstream_BC$discharge <- Upstream_BC$discharge * 1000
Downstream_BC <- read.table(file.path(path_calibration[i], "model_mage", "smooth_bed.LIM"),
  header = FALSE,
  skip = 4,
  col.names = c("time", "stage")
)
Downstream_BC$stage <- Downstream_BC$stage * 1000
# Time assigned for calibration (user defined)
time_observations <- seq(0.95, 3.95, 1) * 60

plot_upstream <- ggplot(Upstream_BC, aes(x = time, y = discharge)) +
  geom_line() +
  geom_vline(xintercept = time_observations, col = "red", linetype = "dashed") +
  annotate("point", x = time_observations[3], y = Upstream_BC$discharge[5], colour = "blue", size = 2.5) +
  labs(
    x = "Simulation time (minutes)",
    y = "Discharge (L/s)",
    title = "Upstream boundary condition"
  ) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


plot_downstream <- ggplot(Downstream_BC, aes(x = time, y = stage)) +
  geom_line() +
  geom_vline(xintercept = time_observations, col = "red", linetype = "dashed") +
  annotate("point", x = time_observations[3], y = Downstream_BC$stage[5], colour = "blue", size = 2.5) +
  labs(
    x = "Simulation time (minutes)",
    y = "Stage (mm)",
    title = "Downstream boundary condition"
  ) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


plot_upstream / plot_downstream

ggsave(plot_upstream,
  filename = file.path("/home/famendezrios/Documents/These/VSCODE-R/HydroBayes/HydroBayes_git/Case_studies/HHLab/Rmd/upstream_boundary_conditions.png"), dpi = 300, units = "px", width = 1200, height = 900
)
ggsave(plot_downstream,
  filename = file.path("/home/famendezrios/Documents/These/VSCODE-R/HydroBayes/HydroBayes_git/Case_studies/HHLab/Rmd/downstream_boundary_conditions.png"), dpi = 300, units = "px", width = 1200, height = 900
)

# Check the observations data:

plot_upstream_example <- ggplot(Upstream_BC, aes(x = time, y = discharge)) +
  geom_line() +
  geom_vline(xintercept = time_observations[3], col = "red", linetype = "dashed") +
  annotate("point", x = time_observations[3], y = Upstream_BC$discharge[5], colour = "blue", size = 2.5) +
  labs(
    x = "Simulation time (minutes)",
    y = "Discharge (L/s)",
    title = "Upstream boundary condition"
  ) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

# Besides, what if we look all observed data
all_observed_zdx <- WS_profiles$case_15_15[, c("x", "z_mean", "Yu", "z_riverbed")]
all_observed_zdx$id <- "All obs"
all_observed_zdx <- all_observed_zdx %>% select("x", "z_mean", "Yu", "id", "z_riverbed")
colnames(all_observed_zdx) <- c("grid", "variable", "uncertainty", "ID", "riverbed")

# What if we look observed Z(X) for the case Q = 120 L/s used during calibration. In other words, at 0.95 hours = 57 minutes
CalData <- read.table(file.path(path_calibration[i], "CalibrationData.txt"), header = TRUE)

WS_obs_example <- CalData[, c(1, 4, nY[i] + 1 + 3)]
WS_obs_example$id <- "Obs calibration"

colnames(WS_obs_example) <- c("grid", "variable", "uncertainty", "ID")

WS_obs_example[WS_obs_example == -9999] <- NA
WS_obs_example <- na.omit(WS_obs_example)

WS_obs_example$riverbed <- all_observed_zdx$riverbed[match(
  WS_obs_example$grid,
  all_observed_zdx$grid
)]


data_example <- rbind(all_observed_zdx, WS_obs_example)
data_example$variable <- data_example$variable * 1000
data_example$uncertainty <- data_example$uncertainty * 1000
data_example$riverbed <- data_example$riverbed * 1000

# Point de fin à annoter avec légende
highlight_df <- data.frame(
  grid = data_example$grid[which.max(data_example$grid)],
  variable = data_example$variable[which.max(data_example$grid)],
  ID = "Downstream"
)


plot_data_example <-
  ggplot(data_example, aes(x = grid, y = variable, col = ID, ymin = variable - uncertainty, ymax = variable + uncertainty, alpha = ID)) +
  geom_point(size = 1) +
  geom_errorbar(width = 0.1, linewidth = 0.5) +
  geom_point(
    data = highlight_df,
    aes(x = grid, y = variable, size = ID), col = "red", alpha = 1, inherit.aes = FALSE,
  ) +
  geom_line(aes(y = riverbed, linetype = "river bed", alpha = "river bed"), color = "brown") +
  scale_color_manual(
    name = "Observations",
    values = c(
      "All obs" = "gray",
      "Obs calibration" = "blue",
      ""
    ),
    labels = c(
      "All obs" = "All",
      "Obs calibration" = "Calibration"
    )
  ) +
  scale_alpha_manual(
    values = c("All obs" = 0.6, "Obs calibration" = 1, "river bed" = 1),
    guide = "none"
  ) +
  labs(
    x = "Lengthwise position (meters)",
    y = "Stage (mm)",
    title = "Water level observations",
    linetype = "Reference",
    size = "Boundary condition"
  ) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

plot_upstream_example / plot_data_example

ggsave(plot_upstream_example,
  filename = file.path("/home/famendezrios/Documents/These/VSCODE-R/HydroBayes/HydroBayes_git/Case_studies/HHLab/Rmd/extraction_obs_q_t.png"), dpi = 300, units = "px", width = 1200, height = 900
)
ggsave(plot_data_example,
  filename = file.path("/home/famendezrios/Documents/These/VSCODE-R/HydroBayes/HydroBayes_git/Case_studies/HHLab/Rmd/extraction_obs_z_x.png"), dpi = 300, units = "px", width = 1200, height = 900
)
```